Using R to help plan the <br> future of transport

### 2018-11-20 img/script.Rmd
update scripts

`img/ms48155lines-mapview.png`:
```{r mapdeck1, eval = FALSE, echo = FALSE}
library (osmdata)
library (dodgr)
library (mapdeck)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
tok <- Sys.getenv ("MAPBOX_TOKEN")
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
```

@@

```{r}
library (dodgr)
gv <- weight_streetnet (ms$osm_lines, wt_profile = "motorcar")
rrr <- 4
index <- match (ids_ring, gv$way_id)
gv$d_weighted [index] <- gv$d [index] / rrr
index <- match (ids_s, gv$way_id)
gv$d_weighted [index] <- gv$d [index] * rrr
gc <- dodgr_contract_graph (gv)
v <- dodgr_vertices (gc$graph)
pts <- sample (v$id, size = 1000)
message ("aggregating ... ", appendLF = FALSE)
fv <- dodgr_flows_aggregate (gv, from = pts, to = pts,
                             flow = matrix (1, nrow = 1000, ncol = 1000))
message ("done\ndispersing ... ", appendLF = FALSE)
fvd <- disperse_flows (fv, 20) %>%
    merge_directed_flows ()
fv <- merge_directed_flows (fv)
mesage ("done")
```
The two maps of vehicular flows:
```{r}
plot1map (fv)
plot1map (fvd)
```
```{r}
f <- readRDS ("data/dodgr-flows-ms70.Rds") %>%
    merge_directed_flows ()
f <- f [which (f$edge_id %in% fvd$edge_id), ]
indx1 <- match (f$edge_id, fvd$edge_id)
indx2 <- match (fvd$edge_id [indx1], f$edge_id)
f$flow <- f$flow / max (f$flow)
fvd$flow <- fvd$flow / max (fvd$flow)
f$flow <- f$flow * fvd$flow [indx1] [indx2]
plot1map (f)
```

>>>>

`img/ms48155lines-mapview.png`:
```{r mapdeck1, eval = FALSE, echo = FALSE}
library (osmdata)
library (dodgr)
library (mapdeck)
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
```

### timing comparison mapview vs mapdeck

```{r}
library (mapview)
library (mapdeck)
system.time (mapview (ms48155$osm_lines))
loc <- as.numeric (apply (sf::st_coordinates (ms48155$osm_lines), 2, mean) [1:2])
system.time (
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
)
```

@@

```{r}
library (dodgr)
gv <- weight_streetnet (ms$osm_lines, wt_profile = "motorcar")
rrr <- 4
index <- match (ids_ring, gv$way_id)
gv$d_weighted [index] <- gv$d [index] / rrr
index <- match (ids_s, gv$way_id)
gv$d_weighted [index] <- gv$d [index] * rrr
gc <- dodgr_contract_graph (gv)
v <- dodgr_vertices (gc$graph)
pts <- sample (v$id, size = 1000)
message ("aggregating ... ", appendLF = FALSE)
fv <- dodgr_flows_aggregate (gv, from = pts, to = pts,
                             flow = matrix (1, nrow = 1000, ncol = 1000))
message ("done\ndispersing ... ", appendLF = FALSE)
fvd <- disperse_flows (fv, 20) %>%
    merge_directed_flows ()
fv <- merge_directed_flows (fv)
message ("done")
```
The two maps of vehicular flows:
```{r}
plot1map (fv)
plot1map (fvd)
```
```{r}
f <- readRDS ("data/dodgr-flows-ms70.Rds") %>%
    merge_directed_flows ()
f <- f [which (f$edge_id %in% fvd$edge_id), ]
indx1 <- match (f$edge_id, fvd$edge_id)
indx2 <- match (fvd$edge_id [indx1], f$edge_id)
f$flow <- f$flow / max (f$flow)
fvd$flow <- fvd$flow / max (fvd$flow)
f$flow <- f$flow * fvd$flow [indx1] [indx2]
saveRDS (f, file = "dodgr-exposure-ms.Rds")
plot1map (f)
```

### 2018-11-20 ms-meetup-nov18.Rmd
update scripts

---
title: "Using R to help plan the <br> future of transport"
subtitle: "Münster R Users Group"
author: "Mark Padgham <br> Active Transport Futures"
date: "Tuesday 20th November, 2018 "
output:
      xaringan::moon_reader:
            yolo:
                        img: img/fortune-beach.png
                        times: 2
            lib_dir: libs
            css: xaringan-themer.css
            nature:
                  highlightStyle: github
                  highlightLines: true
                  countIncrementalSlides: false
---
```{r load, include = FALSE, eval = TRUE}
library(xaringanthemer)
source ("myxaringantheme.R")
source ("script.R")
promenade <- readRDS ("data/osm-promenade.Rds")
ms48155 <- readRDS ("data/osm-ms-48155.Rds")
net <- f <- readRDS ("data/dodgr-flows-ms.Rds")
net$flow <- NULL
```

@@

- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
- `add_osm_feature` `r icon::fa_long_arrow_alt_right()` .bold[WHAT]
- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[HOW]

@@

```{r promenade-data3, eval = FALSE}
library (osmdata)
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "Promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
library (mapview)
mapview (promenade$osm_lines) #<<
```

@@

```{r prom-sf, message = TRUE}
promenade
```

```{r prom-sf-print}
library (sf)
promenade$osm_lines
```
---

@@

```{r promenade-data4, eval = FALSE}
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "Promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
```

@@

@@

```{r, eval = FALSE}
ms <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    osmdata_sf (quiet = FALSE)
```

@@

For more information, see [the website](https://github.com/ropensci/osmdata)

@@

```{r weight-streetnet-ms2, eval = FALSE, message = TRUE}
ms <- osm_poly2line (ms) # osmdata function
library (dodgr)
net <- weight_streetnet (ms$osm_lines, wt_profile = "bicycle") #<<
```

@@

```{r dodgr-vertices1-fakey, eval = FALSE}
v <- dodgr_vertices (net) #<<
dim (v); head (v)
```
```{r dodgr-vertices1, eval = TRUE, echo = FALSE}
v <- dodgr_vertices (net)
dim (v); head (v, n = 5)
```

- `id` is the unique OSM is of each vertex<br>("node" in OSM terminology)
- most `dodgr` functions work either with<br>coordinates or `id` values

@@

@@

- Those 1,000,000 distances could have been calculated with the google API
- And that would cost `1e6 * 0.004 = $US4,000`
- $US 4,000 for 6 seconds of calculation!
- To date, `dodgr` provides one of the only .bold[free] ways to perform such
  calcualtions, and is also entirely .bold[open source].

@@

```{r dodgr-dists1-copy, eval = FALSE}
system.time (
             d <- dodgr_dists (net, from = pts, to = pts)
             )
```
```{r dodgr-dists1-copy-fakey, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
" 43.847   0.013   6.042")
```
```{r dodgr-dists-result1, eval = FALSE}
class (d); dim (d); range (d); range (d, na.rm = TRUE)
```
```{r dodgr-dists-result1-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] matrix\n",    # explicit print coz otherwise output inserts
         "[1] 1000 1000\n", # blank lines that no longer fit on screen
         "[1] NA NA\n",
         "[1]  0.00000 35.50599")
```
```{r dodgr-dists-result2, eval = FALSE}
round (1000 ^ 2 * length (which (is.na (d))) / length (d))
```
```{r dodgr-dists-result2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] 32709")
```

@@

.fonthuge[
mapdeck <br><br> time!
]

>>>>

---
title: "Using R to help plan the <br> future of transport"
subtitle: "Münster R Users Group"
author: "Mark Padgham <br> Active Transport Futures"
date: "Tuesday 20th November, 2018 "
output:
      xaringan::moon_reader:
            yolo:
                        img: img/fortune-beach.png
                        times: 2
            lib_dir: libs
            css: xaringan-themer.css
            nature:
                  highlightStyle: github
                  highlightLines: true
                  countIncrementalSlides: false
---
```{r load, include = FALSE, eval = TRUE}
library(xaringanthemer)
source ("myxaringantheme.R")
source ("data-script.R")
promenade <- readRDS ("data/osm-promenade.Rds")
ms48155 <- readRDS ("data/osm-ms-48155.Rds")
net <- f <- readRDS ("data/dodgr-flows-ms.Rds")
net$flow <- NULL
```

@@

- `opq` `r icon::fa_long_arrow_alt_right()` .bold[WHERE]
- `add_osm_feature` `r icon::fa_long_arrow_alt_right()` .bold[WHAT]
- `osmdata_sf` `r icon::fa_long_arrow_alt_right()` .bold[HOW]

@@

```{r promenade-data3, eval = FALSE}
library (osmdata)
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
library (mapview)
mapview (promenade$osm_lines) #<<
```

@@

@@

```{r promenade-data4, eval = FALSE}
promenade <- opq ("münster de") %>%
    add_osm_feature (key = "highway") %>%
    add_osm_feature (key = "name", value = "promenade",
                     value_exact = FALSE) %>%
    osmdata_sf ()
```

@@

```{r, eval = FALSE}
system.time (
             mapview (ms48155$osm_lines)
)
```
```{r, echo = FALSE, message = TRUE}
res <- c (3.799, 0.074, 3.906)
names (res) <- c ("user", "system", "elapsed")
res
```

```{r, eval = FALSE}
system.time (
mapdeck (style = 'mapbox://styles/mapbox/dark-v9',
         location = loc,
         zoom = 12) %>%
    add_path (data = ms48155$osm_lines,
              layer_id = "ms",
              stroke_width = 1,
              stroke_colour = "#4444dd",
              stroke_opacity = 80)
)
```
```{r, echo = FALSE, message = TRUE}
res <- c (0.222, 0.000, 0.237)
names (res) <- c ("user", "system", "elapsed")
res
```

@@

```{r, eval = FALSE}
ms <- opq ("münster de") %>%               # WHERE
    add_osm_feature (key = "highway") %>%  # WHAT
    osmdata_sf (quiet = FALSE)             # HOW
```

@@

```{r, eval = FALSE}
ms <- opq ("münster de") %>%
    add_osm_feature (key = "building") %>%
    osmdata_sf (quiet = FALSE)
```
- For more information, see [the website](https://github.com/ropensci/osmdata)

@@

@@

@@

```{r dodgr-dists1-copy, eval = FALSE}
system.time (
             d <- dodgr_dists (net, from = pts, to = pts) #<<
             )
```
```{r dodgr-dists1-copy-fakey, echo = FALSE, message = TRUE}
message ("   user  system elapsed\n",
" 43.847   0.013   6.042")
```

@@

```{r dodgr-dists-result1, eval = FALSE}
class (d); dim (d); range (d); range (d, na.rm = TRUE)
```
```{r dodgr-dists-result1-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] matrix\n",    # explicit print coz otherwise output inserts
         "[1] 1000 1000\n", # blank lines that no longer fit on screen
         "[1] NA NA\n",
         "[1] 0.00000 35.50599")
```
```{r dodgr-dists-result2, eval = FALSE}
round (1000 ^ 2 * length (which (is.na (d))) / length (d))
```
```{r dodgr-dists-result2-fakey, eval = TRUE, echo = FALSE, message = TRUE}
message ("[1] 32709")
```

@@

- Those 1,000,000 distances could have been calculated with the google API
- And that would cost `1e6 * 0.004 = $US4,000`
- $US 4,000 for 6 seconds of calculation!
- To date, `dodgr` provides one of the only .bold[free] ways to perform such
  calculations, and is also entirely .bold[open source].

@@

.fonthuge[
mapdeck <br> time!<br>
`r icon::fa_uber(animate = "spin")`
]
---
